---
title: "Functionally enriched epigenetic clocks reveal tissue-specific discordant aging patterns in individuals with cancer"
author: "Main Figures"
date: ""
link-citations: true
output:
  pdf_document:
    toc: false
    latex_engine: xelatex
header-includes:
  \usepackage{fontspec}
  \usepackage[utf8]{inputenc}
  \setmainfont{Open Sans}
  \usepackage{pdfpages}
  \usepackage{graphicx}
  \newfontfamily\boldf{Open Sans Bold}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)

# libraries
library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(RColorBrewer)
library(ggtext)
library(patchwork)
library(here)
library(emojifont)
library(Cairo)

# colours
cols <- MetBrewer::met.brewer("Hiroshige", n = 12)
cols1 <- MetBrewer::met.brewer("Hokusai1", n = 8)
```

\newcommand\invisiblesection[1]{%
  \refstepcounter{section}%
  \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
  \sectionmark{#1}}
  
\newpage

\invisiblesection{Figure 1. Development of cell type-specific functional DNA methylation (DNAme) aging clocks. }

```{r figure1, eval=F}

# Generate panels for figure 1

# b: Overlaps -----------------
# Load and format table
dat <- readxl::read_xlsx(here("0-data/Table-S7.xlsx")) |>
  dplyr::select(-c("Description")) |>
  tidyr::pivot_longer(-c("Set"),
                      names_to = "Subset",
                      values_to = "incl") |>
  tibble::rowid_to_column("order") |>
  dplyr::mutate(incl2 = ifelse(incl == "x" & !is.na(incl), "\u2714",
                               ifelse(incl == "-" & !is.na(incl), "\u2718", NA)),
                group = case_when(grepl("^AGE", Set) ~ "<b>AGE</b>",
                                  grepl("^PCGT", Set) ~ "<b>PCGT</b> sites",
                                  grepl("^SEN", Set) ~ "<b>SEN</b>escence",
                                  grepl("^PRO", Set) ~ "<b>PRO</b>liferation"),
                Subset = factor(Subset, levels = c("Aging (all cells)",
                                                   "Aging (epithelial cells only)",
                                                   "Aging (immune cells only)",
                                                   "PCGT sites",
                                                   "Proliferation-associated CpGs",
                                                   "Proliferation-associated CpGs (age is correlated positively)",
                                                   "Proliferation-associated CpGs (age is correlated negatively)",
                                                   "Senescence-associated CpGs")),
                set = ifelse(grepl("_neg", Set),
                             gsub("_neg", "-", Set),
                             ifelse(grepl("_pos", Set),
                                    gsub("_pos", "+", Set),
                                    paste0(Set))))

b <- dat |>
  ggplot(aes(x = Subset,
             y = forcats::fct_reorder(set, order),
             shape = incl2,
             colour = incl2)) +
  geom_point(size = 3) +
  facet_grid(~group, scales = "free_x",
             space = "free") +
  scale_shape_identity() +
  scale_colour_manual(values = cols[c(11, 1)]) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "top",
        strip.text = element_markdown(hjust = 0),
        panel.grid.major.x = element_blank(),legend.title = element_blank(),
        # panel.grid.major.y = element_blank()
  ) +
  labs(x = "", y = "") +
  coord_flip() +
  scale_x_discrete(limits = rev(levels(dat$Subset)))


grDevices::cairo_pdf(here("4-display-elements/panels/1b.pdf"), width = 13, height = 3.5)
print(b)
dev.off()
```

\begin{center}
\includegraphics[trim=0 0cm 0 0, clip,width=0.8\textwidth]{./main-figure-files/fig1.png}
\end{center}

\begingroup \boldf Figure 1. Development of cell type-specific functional DNA methylation (DNAme) aging clocks. \endgroup

\newpage

\invisiblesection{Figure 2. Initial assessment of cell type-specific functional DNA methylation (DNAme) aging clocks.}

```{r figure2, eval = F}
# a Age specific signatures ------------
load(here("0-data/data_assessment.Rdata"))

a <- data |>
  dplyr::filter(group == "current cancer;surrogate tissue" & type == "Control" & ((celltype == "blood sample" & ic > 0.9) | (celltype != "blood sample" & ic < 0.2))) |>
  dplyr::mutate(menopause = factor(menopause, levels = c("Pre", "Post"))) |>
  tidyr::pivot_longer(cols = c("idx_AGE", "idx_AGEepi", "idx_AGEimm"),
                      names_to = "index",
                      values_to = "value") |>
  dplyr::mutate(index = gsub("idx_", "", index)) |>
  ggplot(aes(x = age,
             y = value,
             colour = celltype)) +
  geom_point(aes(shape = menopause), size = 0.9) +
  facet_wrap(celltype~index, scales = "free") +
  geom_smooth(method = "lm") +
  ggpubr::stat_cor(label.y.npc = "top", inherit.aes = F, label.x.npc= "left",
                   aes(x = age, y = value),
                   method = "spearman") +
  scale_colour_manual(values = cols[c(1, 3, 8)]) +
  scale_shape_manual(values = c(19, 5),
                     name = "Menopausal status") +
  theme_bw() +
  theme(legend.position = "top",
        aspect.ratio = 0.7) +
  labs(x = "chronological age",
       y = "score") +
  guides(colour = "none")

pdf(here("4-display-elements/panels/2a.pdf"), width = 10, height = 6.5)
print(a)
dev.off()

# Source data
source_2a <- ggplot_build(a)$data[[1]]
write.csv(source_2a, file = "4-display-elements/source-data/2a.csv",row.names = F)


# b: Co-correlation --------
tmp <- data |>
  dplyr::filter(group == "current cancer;surrogate tissue" & type == "Control" & ((celltype == "blood sample" & ic > 0.9) | (celltype != "blood sample" & ic < 0.2))) |>
  dplyr::select(idx_AGE:idx_SENpure)
mat <- cor(tmp, method = "spearman")
colnames(mat) <- gsub("idx_", "", colnames(mat))
colnames(mat) <- gsub("_pos", "+", colnames(mat))
colnames(mat) <- gsub("_neg", "-", colnames(mat))
rownames(mat) <- gsub("idx_", "", rownames(mat))
rownames(mat) <- gsub("_pos", "+", rownames(mat))
rownames(mat) <- gsub("_neg", "-", rownames(mat))

b <- Heatmap(mat,
        col = circlize::colorRamp2(breaks = c(-1, -0.75, -0.5, -0.25,
                                              0, 0.25, 0.5, 0.75, 1),
                                   colors = rev(brewer.pal(name = "RdBu", n = 9))),
        name = "Correlation",
        row_names_gp = grid::gpar(fontsize = 10),
        column_names_gp = grid::gpar(fontsize = 10),
        cell_fun = function(j, i, x, y, w, h, fill) {
        if(i >= j) {
            grid.rect(x, y, w, h, gp = gpar(fill = fill, col = fill))
        }
    })

pdf(here("4-display-elements/panels/2b.pdf"), width = 7, height = 6)
draw(b)
dev.off()

# Source data
write.csv(b@matrix, file = "4-display-elements/source-data/2b.csv",row.names = F)

# Reprogramming -----------------
load(here("3-evaluation/3-output/data.Rdata"))
c <- data |>
  dplyr::filter(group == "OSK reprogramming" & celltype == "fibroblast" & accession != "NA") |>
  tidyr::pivot_longer(c(idx_AGE:idx_SENpure),
                         names_to = "index",
                         values_to = "value") |>
  dplyr::mutate(index = gsub("idx_", "", index),
                indexgrp1 = case_when(grepl("^AGE", index) ~ "AGE",
                                     grepl("^PCGT", index) ~ "PCGT",
                                     grepl("^SEN", index) ~ "SEN",
                                     grepl("^PRO", index) & grepl("_neg", index) ~ "PRO (reduction)",
                                     grepl("^PRO", index) & (grepl("_pos", index)|grepl("PROnonage", index)) ~ "PRO (increase)"),
                indexgrp2 = case_when(grepl("gen", index) | index == "AGE" ~ "general",
                                      grepl("epi", index) ~ "epithelial",
                                      grepl("imm", index) ~ "immune",
                                      grepl("pure", index) ~ "pure",
                                      grepl("nonage", index) ~ "nonage"),
                indexgrp2 = factor(indexgrp2, levels = c("general", "epithelial", "immune", "pure", "nonage")),
                index = gsub("_neg", "-", index),
                index = gsub("_pos", "+", index)) |>
  ggplot(aes(x = as.numeric(t),
            y = value,
            colour = index)) +
  geom_point(size = 0.75,
             alpha = 0.9) +
  geom_smooth(alpha = 0.2) +
  geom_text(aes(x = Inf, y = -Inf,
                label = index,
                fontface = "bold"),
                hjust = 1.05, vjust = -7,
            check_overlap = T) +
  # facet_grid(indexgrp1 ~ indexgrp2,
  #            scales = "free", switch = "y") +
  ggh4x::facet_grid2(indexgrp1 ~ indexgrp2,
                     scales = 'free', independent = 'y',
                     switch = "y") +
  theme_bw() +
  theme(legend.position = "none",
        aspect.ratio = 0.7,
        axis.text = element_text(size = 6)) +
  labs(x = "time of reprogramming (days)",
       y = "score") +
  scale_colour_manual(values = c(rep(cols[12], 3),
                                 rep(cols[8], 5),
                                 rep(cols[3], 9),
                                 rep(cols[1], 5)))


pdf(here("4-display-elements/panels/2c.pdf"), width = 9, height = 13)
print(c)
dev.off()

# Source data
source_2c <- ggplot_build(c)$data[[1]]
write.csv(source_2c, file = "4-display-elements/source-data/2c.csv",row.names = F)
```

\begin{center}
\includegraphics[trim=0 0cm 0 0, clip,width=0.8\textwidth]{./main-figure-files/fig2.png}
\end{center}

\begingroup \boldf Figure 2. Initial assessment of cell type-specific functional DNA methylation (DNAme) aging clocks. \endgroup


\invisiblesection{Figure 3. Signature association with physiology and disease.}

```{r figure3, eval=F}
# load dataframes
load("3-evaluation/4-output/df.Rdata")

# combine them
dat <- rbind(dat_osk1, dat_prol, dat_sen, dat_age_cerv, dat_age_buccal, dat_age_blood,
             dat_mort, dat_mort_c, dat_mort_o,
             dat_cur, dat_cur_adj, dat_prog, 
             dat_future, dat_hpv, dat_smk, dat_riskred_brca, dat_riskred_wt,
             dat_cur_sur, 
             # dat_future_sur,
             dat_mort_blood, dat_mort_c_blood)

dat <- dat |>
  dplyr::mutate(p = p.adjust(p, method = 'fdr'))

# brng into the right shape for plotting as a matrix
tmp1 <- dat |>
  dplyr::mutate(index = gsub("idx_|_adj", "", index),
                index = gsub("_neg", "-", index),
                index = gsub("_pos", "+", index),
                index = gsub("hannum", "Hannum clock", index),
                index = gsub("horvath", "Horvath clock", index),
                index = gsub("phenoage", "PhenoAge", index)) |>
  tidyr::pivot_wider(id_cols = c(group, celltype, type, n_control, n_case),
                     names_from = "index",
                     values_from = "auc") |>
  dplyr::mutate(label = case_when(group == "current cancer" & type == "Cancer" ~ paste0("cancer (", celltype, ")"),
                                  group == "current cancer" & grepl("CIN", type) ~ paste0(type, " (cervix)"),
                                  group == "current cancer" & type != "Cancer" & !grepl("CIN", type) ~ paste0(tolower(type), " (breast tissue)"),
                                  group == "response" ~ "non-complete neo-adjuvant chemotherapy\nresponse",
                                  type == "HPV positive" ~ paste0("HPV (cervical sample)"),
                                  type == "Mifepristone (BRCA)" ~ paste0("mifepristone (BRCAmut breast)"),
                                  type == "Mifepristone (WT)" ~ paste0("mifepristone (WT breast)"),
                                  grepl("future", group) & !grepl("surrogate", group) ~ "cyt normal, future CIN3+ (cervix)",
                                  group == "proliferation" ~ paste0(substr(tolower(type), 1, nchar(type)-1), ", ", celltype, ")"),
                                  group == "senescence" ~ paste0(tolower(type), " (", celltype, ")"),
                                  grepl("surrogate", group) & grepl("current", group) ~ paste0(tolower(type), " (", celltype, ")"),
                                  # grepl("surrogate", group) & grepl("future", group) ~ paste0(tolower(type), " (future, ", celltype, ")"),
                                  grepl("smoker", type, ignore.case = T) ~ paste0(tolower(type), " (buccal sample)"),
                                  grepl("mortality;all-cause mortality$", group) & celltype == "buccal sample" ~ "all-cause 10-year mortality (buccal sample)",
                                  grepl("mortality;all-cause mortality$", group) & celltype == "blood sample" ~ "all-cause 10-year mortality (blood sample)",
                                  grepl("mortality;cancer", group) & celltype == "buccal sample" ~ "cancer-specific 10-year mortality (buccal sample)",
                                  grepl("mortality;cancer", group) & celltype == "blood sample" ~ "cancer-specific 10-year mortality (blood sample)",
                                  grepl("mortality;other", group) ~ "non-cancer 10-year mortality (buccal sample)",
                                  TRUE ~ tolower(type))) |>
  # dplyr::mutate(label = paste0(label, ", n = ", n)) |>
  dplyr::mutate(group2 = case_when(group == "current cancer" & grepl("Normal|CIN", type) ~ "pre-cancer or adjacent",
                                   group == "current cancer" & !grepl("Normal|CIN", type) ~ "current cancer",
                                   group == "future cancer;tissue at risk" ~ "future cancer",
                                   grepl("surrogate", group) & grepl("current", group) ~ "surrogate tissue (current cancer)",
                                   # grepl("surrogate", group) & grepl("future", group) ~ "surrogate tissue (future cancer)",
                                   grepl("mortality", group) ~ "mortality",
                                   group == "response" ~ "non-complete neo-adjuvant chemotherapy response",
                                   TRUE ~ paste0(group)),
                group3 = case_when(group2 %in% c("OSK reprogramming", "senescence", "proliferation", "normal aging") ~ "reprogramming, aging, and senescence",
                                   group2 %in% c("current cancer", "pre-cancer or adjacent", "future cancer") ~ "cancer",
                                   grepl("risk", group2) ~ "exposures",
                                   group == "response" ~ "cancer",
                                   grepl("surrogate", group2) ~ "surrogate tissue (cancer)",
                                   grepl("mortality", group2) ~ "mortality",
                                   TRUE ~ paste0(group2)),
                ref = case_when(group %in% c("OSK reprogramming", "proliferation") ~ "Fibroblast",
                                group == "senescence" & grepl("fibroblast", celltype) ~ stringr::str_to_sentence(celltype),
                                group == "senescence" & grepl("endothelial", celltype) ~ "HUVEC",
                                group == "senescence" & grepl("mesenchymal", celltype) ~ "MSC",
                                group == "normal aging" ~ "Pre-menopausal sample",
                                 TRUE ~ "Ref"),
                comp = case_when(group == "OSK reprogramming" ~ "iPSC/ESC",
                                 group == "proliferation" ~ "Serum withdrawal",
                                 group == "senescence" & grepl("fibroblast", celltype) ~ type,
                                 group == "normal aging" ~ "Post-menopausal sample",
                                 TRUE ~ "Comp"))
mat1 <- tmp1[,6:34]
rownames(mat1) <- tmp1$label

# Transform sigmat
tmp2 <- dat |>
  dplyr::mutate(index = gsub("idx_|_adj", "", index),
                index = gsub("_neg", "-", index),
                index = gsub("_pos", "+", index),
                index = gsub("hannum", "Hannum clock", index),
                index = gsub("horvath", "Horvath clock", index),
                index = gsub("phenoage", "PhenoAge", index)) |>
  dplyr::rowwise() |>
  dplyr::mutate(aucsig = ifelse(all(c(auc, cilo, cihi) > 0.5) | all(c(auc, cilo, cihi) < 0.5),
                                gtools::stars.pval(p),
                                ""),
                aucsig = ifelse(grepl("*", aucsig, fixed = T),
                                "*",
                                ifelse(grepl(".", aucsig, fixed = T),
                                       "", aucsig))) |>
  ungroup() |>
  tidyr::pivot_wider(id_cols = c(group, celltype, type, n_control, n_case),
                     names_from = "index",
                     values_from = "aucsig")
mat2 <- tmp2[,6:34]

cols <- MetBrewer::met.brewer("Hiroshige", n = 12)
cols1 <- MetBrewer::met.brewer("Hokusai1", n = 9)

hl = rowAnnotation(
  group = factor(tmp1$group3, levels = c("reprogramming, aging, and senescence",
                                         "mortality",
                                         "cancer",
                                         "exposures",
                                         "surrogate tissue (cancer)")),
  type = factor(tmp1$group2, levels = c("OSK reprogramming",
                                                          "senescence",
                                                          "proliferation",
                                                          "normal aging",
                                        "mortality",
                                                          "current cancer",
                                        "non-complete neo-adjuvant chemotherapy response",
                                                          "pre-cancer or adjacent",
                                                          "future cancer",
                                                          "risk-increasing exposure",
                                                          "risk-reducing exposure",
                                        "surrogate tissue (current cancer)")),
                   col = list(group = c("reprogramming, aging, and senescence" = cols1[1],
                                        "mortality" = cols1[3],
                                        "cancer" = cols1[5],
                                        "exposures" = cols1[7],
                                        "surrogate tissue (cancer)" = cols1[8]),
                              type = c("OSK reprogramming" = cols[1],
                                        "senescence" = cols[2],
                                        "proliferation" = cols[3],
                                        "normal aging" = cols[4],
                                       "mortality" = cols[5],
                                        "current cancer" = cols[6],
                                       "non-complete neo-adjuvant chemotherapy response" = cols[7],
                                        "pre-cancer or adjacent" = cols[8],
                                        "future cancer" = cols[9],
                                        "risk-increasing exposure" = cols[10],
                                        "risk-reducing exposure" = cols[11],
                                       "surrogate tissue (current cancer)" = cols[12]),
                                  simple_anno_size = unit(0.5, "cm"),
                                  show_annotation_name = T)
                )

# anno_n
hr = rowAnnotation(
  # `Reference` = anno_text(tmp1$ref,
  #                                          show_name = T,
  #                                          gp = grid::gpar(fontsize = 8)),
                   `Reference (n)` = anno_numeric(tmp1$n_control, width = unit(0.3, "in"),
                                               align_to = "right",
                                          labels_gp = grid::gpar(fontsize = 8),rg = c(0, 400),
                                           bg_gp = gpar(fill = "grey80", col = NA)),
                   `Comparison (n)` = anno_numeric(tmp1$n_case,  width = unit(0.3, "in"),
                                            rg = c(0, 400),
                                          labels_gp = grid::gpar(fontsize = 8),
                                           bg_gp = gpar(fill = "grey70", col = NA)),
                   # `Comparison` = anno_text(tmp1$comp,
                   #                         show_name = T,
                   #                         gp = grid::gpar(fontsize = 8)),
                   annotation_name_rot = 90)

map <- Heatmap(as.matrix(mat1),
        cluster_rows = T,
        cluster_columns = T,
        show_row_names = T,
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(mat2[i, j], x, y, gp = gpar(fontsize = 10))
        },
        col = circlize::colorRamp2(breaks = c(0, 0.125, 0.25, 0.375,
                                              0.5, 0.625, 0.75, 0.875, 1),
                                   colors = rev(brewer.pal(name = "RdBu", n = 9))),
        name = "AUC values",
        row_names_gp = grid::gpar(fontsize = 8),
        column_names_gp = grid::gpar(fontsize = 9),
        row_title_gp = grid::gpar(fontsize = 10),
        row_split = factor(tmp1$group2, levels = c("OSK reprogramming",
                                                   "senescence",
                                                   "proliferation",
                                                   "normal aging",
                                                   "mortality",
                                                    "current cancer",
                                                   "non-complete neo-adjuvant chemotherapy response",
                                                      "pre-cancer or adjacent",
                                                   "future cancer",
                                                      "risk-increasing exposure",
                                                      "risk-reducing exposure",
                                                   "surrogate tissue (current cancer)")),
        row_gap = unit(c(1, 1, 1, 5, 5, 1, 1, 1, 5, 1, 5, 5), "mm"),
        show_row_dend = F,
        row_title = NULL,
        border = T,
        left_annotation = hl,
        right_annotation = hr,
        cluster_row_slices = F)

grDevices::cairo_pdf(here("4-display-elements/panels/heatmap.pdf"), width = 11, height = 10)
draw(map)
dev.off()

# heatmap is annotated in illustrator (labels)]

# Source data
write.csv(map@matrix, file = "4-display-elements/source-data/3.csv",row.names = F)
```

\begin{center}
\includegraphics[trim=0 0 0 0, clip,width=0.8\textwidth]{./main-figure-files/fig3.png}
\end{center}

\begingroup \boldf Figure 3. Signature association with physiology and disease. \endgroup

\invisiblesection{Figure 4. Mouse model of carcinogen exposure and prevention with multi-tissue methylation analysis.}

```{r figure4, eval = F}
# Load mouse data
load("0-data/dat_assessment_mouse.Rdata")

# Prepare pheno for plotting
pheno <- pheno |> 
  # Filter Exp 5 and relevant tissues only
  dplyr::filter(substudy == "Cancer study" & tissue %in% c("Cervix", "Breast", "PBMCs")) |> 
  droplevels() |> 
  # create new type3 variable for plotting              
  dplyr::mutate(type3 = case_when(treatment == "MPA/DMBA- MIF-" & tumour_status == "No tumor" ~ "Control",
                                  treatment == "MPA/DMBA+ MIF-" & tumour_status == "Tumor" ~ "Cancer case",
                                  treatment == "MPA/DMBA+ MIF+" & tumour_status == "No tumor" ~ "Carcinogen exposure with successful preventive measure",
                                  
                                  treatment == "MPA/DMBA+ MIF+" & tumour_status == "Tumor" ~ "Carcinogen exposure with unsuccessful preventive measure",
                                  treatment == "MPA/DMBA- MIF+" & tumour_status == "No tumor" ~ "Preventive measure only"),
  # relevel type3 for right appearance
                type3 = factor(type3, levels = c("Control",
                                                 "Cancer case",
                                                 "Carcinogen exposure with unsuccessful preventive measure",
                                                 "Carcinogen exposure with successful preventive measure",
                                                 "Preventive measure only")),
  # fix tissue
  tissue = case_when(tissue == "Breast" ~ "mammary gland",
                     tissue == "Cervix" ~ "cervix",
                     tissue == "PBMCs" ~ "PBMCs"),
  tissue = factor(tissue, levels = c("mammary gland", "cervix", "PBMCs"))) |> 
  dplyr::rename(mPCGTgen = idx_mPCGTgen,
                mPCGTnonage = idx_mPCGTnonage) |> 
  tidyr::pivot_longer(cols = c("mPCGTgen", "mPCGTnonage"),
                      names_to = "index",
                      values_to = "value") 

theme_set(
  theme_bw() +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = 'right')
)

plot <- pheno |>
  ggplot(
    aes(x=type3, y=value)
    ) +
  ggbeeswarm::geom_beeswarm(
    aes(colour = type3,
        shape = tumour_status),
    size = 2
    ) +
  geom_boxplot(
    aes(fill = type3),
    alpha = 0.7,
    outlier.shape = NA
    ) +
  facet_grid(
    index~tissue,
    switch = "y",
    scales = "free_y"
    ) +
  ggpubr::stat_compare_means(
    label.x.npc = 0.1,
    size = 3
    # comparisons = list(c("Control", "Cancer case"),
    #                    c("Carcinogen exposure with unsuccessful preventive measure", "Carcinogen exposure with successful preventive measure"),
    #                    c("Cancer case", "Carcinogen exposure with successful preventive measure"),
    #                    c("Control", "Preventive measure only")
    #                    )
    # ,
    #                    c("Cancer case", "Carcinogen exposure with successful prevention"),
    #                    c("Control", "Preventive measure only"))
  ) +
  labs(
    x = "",
    y = ""
  ) +
  scale_colour_manual(
    values = cols1[c(8, 1, 4, 6, 7)],
    aesthetics = c("colour", "fill"),
    name = ""
    ) +
  scale_shape_manual(
    values = c(1, 19),
    name = ""
  ) +
  guides(color = guide_legend(nrow = 5),
         shape = guide_legend(nrow = 2)) +
  theme(
    aspect.ratio = 1.3
  )


# plot <- pheno |>
#   ggplot(
#     aes(x=type3, y=value)
#     ) +
#   ggbeeswarm::geom_beeswarm(
#     aes(colour = type3,
#         shape = tumour_status),
#     size = 2
#     ) +
#   geom_boxplot(
#     aes(fill = type3),
#     alpha = 0.7,
#     outlier.shape = NA
#     ) +
#   facet_grid(
#     index~tissue,
#     switch = "y",
#     scales = "free_y"
#     ) +
#   ggpubr::stat_compare_means(
#     ref.group = 'Control',
#     label.x.npc = 0.1,
#     label = 'p.format',
#     label.y.npc = 0.95,
#     size = 2.5
#     # comparisons = list(c("Control", "Cancer case"),
#     #                    c("Carcinogen exposure with unsuccessful preventive measure", "Carcinogen exposure with successful preventive measure"),
#     #                    c("Cancer case", "Carcinogen exposure with successful preventive measure"),
#     #                    c("Control", "Preventive measure only")
#     #                    )
#     # ,
#     #                    c("Cancer case", "Carcinogen exposure with successful prevention"),
#     #                    c("Control", "Preventive measure only"))
#   ) +
#   labs(
#     x = "",
#     y = ""
#   ) +
#   scale_colour_manual(
#     values = cols1[c(8, 1, 4, 6, 7)],
#     aesthetics = c("colour", "fill"),
#     name = ""
#     ) +
#   scale_shape_manual(
#     values = c(1, 19),
#     name = ""
#   ) +
#   guides(color = guide_legend(nrow = 5),
#          shape = guide_legend(nrow = 2)) +
#   theme(
#     legend.position = 'bottom'
#   )
# 
# plot

grDevices::cairo_pdf(file = here("4-display-elements/panels/4b.pdf"),
       width = 13, 
       height = 5)
print(plot)
dev.off()

# Source data
source_4b <- ggplot_build(plot)$data[[1]]
write.csv(source_4b, file = "4-display-elements/source-data/4b.csv",row.names = F)
```


\begin{center}

